Cost- and Deadline-constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds

### 2012-01-18 paper/paper.tex
updated paper with new SPSS

Initially all the
released tasks from all the workflows are submitted to the queue $Q$. Then, they
are inserted into priority queue $P$ based on the priority of workflows. If
there are free VMs available, the tasks from the priority queue are submitted to
the randomly choosen VMs. The process is repeated each time a task finishes.
Using the priority queue guarantees that tasks from lower priority workflows are
always submitted later than the higher priority ones and that the lower priority
tasks can occupy idle VMs. However, as there is no preemption, long-running
low-priority tasks may delay execution of high-priority tasks when such ones
become ready. Fig.~\ref{fig:spss-example} shows the example plan generated using
SPDS algorithm. It can be seen that tasks from lower priority workflows can fill
idle time of provisioned virtual machines.

\begin{figure}[htb] 
\centering
\includegraphics[width=1.0\columnwidth]{figures/spds-gantt}
 \caption{Example of schedule generated using SPDS algorithm: tasks are labelled
 by colors depending on the workflow. }
\label{fig:spss-example}
\end{figure}

@@

@@

@@

@@

\begin{figure}[htb] 
\centering
\includegraphics[width=0.6\columnwidth]{figures/ensemble-pareto}
\caption{Pareto-like distribution of workflow sizes in ensemble. }
\label{fig:ensemble-distribution}
\end{figure}

@@

@@

@@

\begin{figure*}[t] 
\centering
\includegraphics[width=0.19\textwidth]{figures/MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Number of workflows completed for DPDS, AW-DPDS and SPSS
algorithms on ensembles of 100 Pareto-distributed workflows}
\label{fig:number-complete-pareto}
\end{figure*}

@@

\begin{figure}[b!] 
\centering
%\includegraphics[width=0.19\textwidth]{figures/score-MONTAGE-n-1000-8-dagh1-20m0.pdf}
%\includegraphics[width=0.19\textwidth]{figures/score-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/score-LIGO-n-1000-8-dagh1-40m0.pdf}
%\includegraphics[width=0.19\textwidth]{figures/score-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/score-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Comparison of algorithms based on score, which  is computed as
$\sum(1/(p+1))$ where priorities of subsequent workflows are $p=0,1,2,\ldots$ (0 means highest
priority).}
\label{fig:score}
\end{figure}

@@

\begin{figure*}[htb]  
\centering
\includegraphics[width=0.19\textwidth]{figures/size-MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/size-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/size-LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/size-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/size-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
Pareto-distributed workflows. Score is equal to the total work completed,
i.e. it is sum of runtimes (in hours) of tasks of completed workflows.}
\label{fig:total-time}
\end{figure*}

\begin{figure*}[b!] 
\centering
\includegraphics[width=0.19\textwidth]{figures/cost-MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/cost-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/cost-LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/cost-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/cost-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Effective computing cost in \$ per hour calclated by dividing sum of
task runtimes by the total budget. Higher cost results from lower resource utilization. Data based on DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
Pareto-distributed workflows.}
\label{fig:cost}
\end{figure*}

@@

@@

\begin{figure}[htb] 
\centering
\includegraphics[width=0.6\columnwidth]{figures/budget-SIPHT-n-1000-8-dagh5-50h10m0.pdf}
\caption{Comparison of number of workflows completed for a given deadline,
plotted versus different budget amounts.}
\label{fig:}
\end{figure}

@@

>>>>

Initially all the
released tasks from all the workflows are submitted to the queue $Q$. Then, they
are inserted into priority queue $P$ based on the priority of workflows. If
there are free VMs available, the tasks from the priority queue are submitted to
the randomly choosen VMs. The process is repeated each time a task finishes.
Using the priority queue guarantees that tasks from lower priority workflows are
always submitted later than the higher priority ones and that the lower priority
tasks can occupy idle VMs. However, as there is no preemption, long-running
low-priority tasks may delay execution of high-priority tasks when such ones
become ready. Fig.~\ref{fig:spds-example} shows the example plan generated using
SPDS algorithm. It can be seen that tasks from lower priority workflows can fill
idle time of provisioned virtual machines.

\begin{figure}[htb] 
\centering
\includegraphics[width=1.0\columnwidth]{figures/spds-gantt}
 \caption{Example of schedule generated using SPDS algorithm: tasks are labelled
 by colors depending on the workflow. }
\label{fig:spds-example}
\end{figure}

@@

DPDS algorithm does not use any information about workflow structure: it deals
only with scheduling of individual released/ready tasks using their
workflow priorities. If there are no high priority tasks in the ready queue,
then lower priority tasks are submitted to idle resources. This may lead to the
deficiency in the following scenario: if a long-running task of low
priority is submitted to the idle machine and then several high
priority tasks are released to the ready queue, they become delayed by the
execution of the low priority task. This usually does not cause a problem when
the number of tasks and resources is high, but when the execution approaches the
budget and deadline constraint, we have observed that tasks of low-priority
workflows may prevent higher priority workflows from completion. This
observation leads to the improvemed algorithm which we call workflow-aware DPDS
(WA-DPDS).

WA-DPDS extends DPDS by introducing a workflow admission procedure, which is
invoked each time a task of a new workflow is at the head of priority queue,
i.e. just before it is submitted. The procedure estimates if there is enough
budget remaining to admit a new workflow: if not, all its tasks are rejected.
Algorithm~\ref{alg:wa-dpds} estimates the cost of admitting a new workflow and
compares it to the current cost (consumed budget) and remaining budget, taking
into account that started VMs have been already paid for. It adds a small safety
margin choosen to avoid overestimation of a budget.

\begin{algorithm}
\caption{Workflow admission algorithm for WA-DPDS}
\label{alg:wa-dpds}
\begin{algorithmic}[1]
\Require $w$: workflow; $b$: budget; $c$: current cost
\Procedure{Admit}{$w,b,c$}
    \State $r_n\gets b-c$ \Comment{Remaining budget for new VMs}
    \State $r_c\gets remaining\ cost\ on\ running\ VMs$
    \State $r_a\gets remaining\ cost\ of\ admitted\ workflows$
	\State $r_m\gets 0.1$ \Comment{Safety margin}
	\State $r_b\gets r_n+r_c-r_a-r_m$ \Comment{Budget remaining}
	\State $c_w\gets estimate\ cost(w)$
    \If{$c_w<r_b$}
    	\State \textbf{return} $TRUE$
    \Else
    	\State \textbf{return} $FALSE$
	\EndIf    	 
\EndProcedure
\end{algorithmic} 
\end{algorithm}

Admission procedure relies only on the total estimated resource consumption and
compares it to the budget remaining. We found that this estimation is useful not
only to prevent low-priority workflows from delaying ligh-priority ones, but
also in the case of ensembles consisting of workflows of non-uniform sizes, it
allows to reject big and costly workflows that would overrun the budget and
instead schedule some smaller ones to efficiently utilize idle resources. It
would be also possible to extend the admission procedure to check other
constraints, such as whether estimated critical path exceeds the deadline.

@@

\begin{figure}[htb] 
\centering
\includegraphics[width=1.0\columnwidth]{figures/spss-gantt}
 \caption{Example of schedule generated using SPSS algorithm: tasks are labelled
 by colors depending on the workflow. }
\label{fig:spss-example}
\end{figure}

@@

\begin{figure}[tb] 
\centering
\includegraphics[width=0.6\columnwidth]{figures/ensemble-pareto}
\caption{Pareto-like distribution of workflow sizes in ensemble. }
\label{fig:ensemble-distribution}
\end{figure}

@@

\begin{figure*}[t] 
\centering
\includegraphics[width=0.19\textwidth]{figures/pareto-MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Number of workflows completed for DPDS, AW-DPDS and SPSS
algorithms on ensembles of 100 Pareto-distributed workflows}
\label{fig:number-complete-pareto}
\end{figure*}

\begin{figure}[b!] 
\centering
%\includegraphics[width=0.19\textwidth]{figures/pareto-score-MONTAGE-n-1000-8-dagh1-20m0.pdf}
%\includegraphics[width=0.19\textwidth]{figures/pareto-score-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-score-LIGO-n-1000-8-dagh1-40m0.pdf}
%\includegraphics[width=0.19\textwidth]{figures/pareto-score-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-score-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Comparison of algorithms based on score, which  is computed as
$\sum(1/(p+1))$ where priorities of subsequent workflows are $p=0,1,2,\ldots$ (0 means highest
priority).}
\label{fig:score}
\end{figure}

@@

\begin{figure*}[t]  
\centering
\includegraphics[width=0.19\textwidth]{figures/pareto-size-MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-size-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-size-LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-size-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-size-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{ Total work completed expressed as the sum of runtimes (in hours) of
tasks of completed workflows.Data computed for DPDS, AW-DPDS and SPSS
algorithms on ensembles of 100 Pareto-distributed workflows}
\label{fig:total-time}
\end{figure*}

@@

We show the results of three algorithms (TODO: ADD MORE): DPDS, WA-DPDS and
SPSS. These tests were run with maximum autoscaling factor set to 1.0 for
DPDS and WA-DPDS. After experimenting with DPDS and WA-DPDS we found that due to
high parallelism of workflows the resource utilization remains high most of the time, 
which leads to rapid provisioning of new resources at the beginning of ensemble
execution. This in turn results in less efficient total utilization due to high
level of parallelism. Setting the scaling limit to 1.0, ensures that the shape
of resource provisioning plan is closest to a rectangle (see
Fig.~\ref{fig:spds-example}) and the VM termination procedure
(Algorithm~\ref{alg:prov}) allows maximum utilization of VMs before shutdown.

@@

@@

@@

\begin{figure*}[htb] 
\centering
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Effective computing cost in \$ per hour calclated by dividing sum of
task runtimes by the total budget. Higher cost results from lower resource utilization. Data based on DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
Pareto-distributed workflows.}
\label{fig:cost}
\end{figure*}

@@

\begin{figure}[htb] 
\centering
\includegraphics[width=0.6\columnwidth]{figures/pareto-budget-SIPHT-n-1000-8-dagh5-50h10m0.pdf}
\caption{Comparison of number of workflows completed for a given deadline,
plotted versus different budget amounts.}
\label{fig:budgets}
\end{figure}

@@

Fig.~\ref{fig:budgets} whows the same data for SIPHT ensemble from another
perspective. It helps answer the question: given a deadline, how many workflows
can be finished when we increse the budget. In Fig.~\ref{fig:budgets} the growth
is super-linear, what results from the distribution of workflow sizes in the
ensemble (see Fig.~\ref{fig:ensemble-distribution}). For ensemble of equal size
workflows (see Fig. TODO: ADD FIGURE FROM CONSTANT DISTRIBUTION) the growth is
nearly linear. It can be also observed that staitc algorithms give better
results for smaller budgets, while dynamic algorithms for larger ones.

@@

Summary:

 - new interesting problem was addressed and analyzed

 - dynamic and static strtegies were evaluated

 - new algorithms were developed

 - simulation studies were done

Lessons learned:

 - pros and cons of dynamic/static methods

 - problems faced and solved

 - observations about parallelization level, shape of the plan

Future work:

 - deal with uncertaineties of predictions

 - deal with heterogeneity and multiple clouds (ref to Gideon's and Jens'
 experiments on EC2 and FutureGrid)

 - add data transfers to cost models (ref to paper on data sharing)
